The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation

### 2019-12-02 writing/manuscript.Rmd
Continnued with results sections

## Crossover rate

@@

@@

@@

>>>>

We first tested whether we were succsessful in reimplementing the model made by @Wittmann2017. We too found that with no linkage (i.e. cross-over happens at every loci) both the proportion of fixed loci (p = 2.52e-07) and standard deviation of loci frequencies (p = 4.63e-07) are lower when dominance is 0.8 compared to when dominance is 0.5.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each value of the epistatis parameter. \\label{epi_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/cross_over"))

# load all files and filter only the final generation
all_tbl <- list.files(pattern = "c_0.5") %>% 
  map_df(~read_plus(.))

# plot the density distribution
p1 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6)],
                          # legend name
                          name = "Dominance",
                          # values tested
                          labels = c("0.5", "0.8")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

```{r, fig.width = 10}
# Small fig.width
knitr::include_graphics("D05C05.png")
knitr::include_graphics("D08C05.png")
```

Then, we started playing with different parameters.
## Crossover rate

Crossover rate didn't seem to have an effect at all in terms of standard deviation (p = 0.123) and proportion of fixed loci (p = 0.462). Our hypothesis is that crossover didn't have an effect because two opposing forces created by it canceled out. First of all, higher crossover rates creates more genetic variation in the population "helping" selection but on the other hand it also creates stochasticity since a selected high fitness individual can pass on an unfit chromosome after crossover by chance.

When there was no crossover and so linkage was maximized there was a much bigger variation in the proportion of fixed loci in the last generation. This could be because since in this case the only way for an individual to change was through mutations, which is pretty infrequent, if the starting population contained a lot of individuals with more of one allele than the other many loci went to fixation while in a starting population with individuals whose genomes contained 50%-50% of the two alleles balancing polymorphism could be maintained more successfully.

```{r, message = FALSE, include=FALSE}
setwd(here::here("results/cross_over"))
values = c("c_0_", "c_0.001", "c_0.01_", "c_0.05","c_0.1", "d_0.5_y_1_c_0.5_")
b <- do_analysis(values, "Dominance", "c_", "b3")
```
```{r, echo = FALSE}
b
```

Since the crossover rate didn't have an effect on whether or not balanced polymorphism is achieved, our tests with different dominance values resulted in the same dynamics as when we tested with crossover rate 0.5.

@@

Under default crossover rate (0.05) epistasis didn't have an effect (standard deviation p = 0.986, proportion of fixed alleles p = 0.613).

@@

When the generations per season was 10, the standard deviation was significantly lower than for longer generation times per season (p = 0.00161). We hypothise that with more generations per season loci were under selection for a longer period of time and thus were more likely to go to fixation at one point.

@@

```{r, fig.width = 10}
# Small fig.width
knitr::include_graphics("G10.png")
knitr::include_graphics("G50.png")
```

Population size had the biggest effect by far (standard deviation p = 1.51e-10, proportion of fixed loci p = 1.63e-10). Purely by changing population size from 100 to 1000 average proportion of fixed loci decreased from 0.9 to 0.3. This is probably due to the effect of random genetic drift; when the population is small alleles go to fixation due to the large effect of drift.

```{r}
setwd(here::here("results/pop_size"))
values = c("Ps_100_", "Ps_500", "Ps_1000")
b <- do_analysis(values, "Generations per semester", "Ps_", "b3")
```
```{r, echo = FALSE}
b
```

Interestingly, having a season that has more generations than in the other didn't result in a higher proportion of fixed alleles as one whould have expected (p = 0.289). Rather, while alleles fixed by reaching frequencies of 0 or 1 in the case where both seasons had equal number of generations (seasonal balance = 2) when there were more generations in one more alleles had frequencies of 1 than 0 but the overall proportion stayed the same.

```{r, fig.height = 3, fig.width = 5, fig.cap = 'A density plot of the allele frequencies at the final generation for five replicates of each number of generations per year. \\label{gen_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/seasonal_balance"))

# plot the density distribution
p3 <- ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
        geom_density(alpha = 0.2) + 
        scale_fill_manual(values = c(rep("#e25dbb", 5), 
                                     rep("#6cdf4c", 5),
                                     rep("#258bd2", 5)),
                          breaks = unique(all_tbl$filename)[c(1,6,11)],
                          # legend name
                          name = "Generations\nper year",
                          # values tested
                          labels = c("1.25", "1.5", "2")) +
        ylab("Density") + 
        xlab("Allele Frequency") +
        theme_classic()

print(p3)

## Mutational interaction

Since crossover rate didn't have an effect we decided to test the scenario when loci have a certain interaction between them. In real genomes mutations often have different fitness effects depending on the genetic background i.e. the presence of other mutations. Thus, linkage can be advantageous if specific combinations of mutations are better kept together because of their reinforcing effect on each other. In this simulation we kept the same fitness function as before but also rewarded sequences of the benefitial allele. For example, in the summer an individual with chromosome 010101 would have a fitness of 3 while an individual with chromosome 011000 would have a fitness of 2 + 2 (for the longest sequence in the genome) which would then outcompete the former individual eventhough the total number of 1's is less.

```{r}
setwd(here::here("results/mut_interaction"))
values = c("c_0.5_new", "c_0.05_new", "c_0.5_num", "c_0.05_num")
b <- do_analysis(values, "Generations per semester", "none", "b3")
```
